---
title: "Unemployment in America Per US State"
name: "Fabian Berger"
date: "23. Februar 2023"
output: 
  html_document: 
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Inspection of the data

This dataset contains relevant statistics and employment rates per US state since 1976.

Data is availabe at: [Kaggle](https://www.kaggle.com/datasets/justin2028/unemployment-in-america-per-us-state?resource=download)

According to the description on the website, the dataset from *Kaggle* does not contain any missing values. The data was briefly checked for any missing and infinitive values.

```{r loading}
#load packages
library(tidyverse)
library(janitor)
library(ggridges)
library(viridis)
library(hrbrthemes)
library(ggiraph)
library(usmap)
library(psych)
library(kableExtra)

library(cluster)
library(factoextra)
library(proxyC)

library(broom)

#load data from 
data <- read_csv('Unemp_usa.csv')



glimpse(data)

#clean names/whitespaces
data_clean <- clean_names(data)

#check for any NA and infite values
check_na_inf <- apply(data_clean, 2, function(x) any(is.na(x) | is.infinite(x)))
print(check_na_inf)


```

```{r}
#write a histogram function
hist_func <- function(data, var,z,title) {
  ggplot(data = data, mapping = 
           aes(x = var)) +
    geom_histogram(bins = z) +
    scale_x_continuous(labels = scales::comma) +
    theme_minimal() +
    labs(title = title)
}

#test function x=data, y=variable, z=bins, title=title
hist_func(data_clean, var = data_clean$total_civilian_non_institutional_population_in_state_area, 150, "Population in States")


  
```


## EDA

```{r}
data_grouped_by_year_state <- data_clean %>% 
  group_by(year,state_area) %>%
  summarise(`Total non-institutional Population (USA)` = sum(total_civilian_non_institutional_population_in_state_area),
            `Total Civilian Labor Force (USA)` = sum(total_civilian_labor_force_in_state_area),
            `Total Employment in USA` = sum(total_employment_in_state_area)
            )

data_grouped_by_year_state <- data_grouped_by_year_state %>% 
  mutate(`Percentage of Civilian Labor Force (USA)` =  (`Total Civilian Labor Force (USA)`/`Total non-institutional Population (USA)`)*100,
         `Percentage of Employment in USA` = (`Total Employment in USA`/`Total non-institutional Population (USA)`)*100,
         `Percentage of Unemployment in USA` = ((`Total Civilian Labor Force (USA)`-`Total Employment in USA`)/`Total non-institutional Population (USA)`)*100
  )




            
data_nat_year <- data_clean %>% 
  group_by(year) %>% 
  summarise(`Total non-institutional Population (USA)` = sum(total_civilian_non_institutional_population_in_state_area),
            `Total Civilian Labor Force (USA)` = sum(total_civilian_labor_force_in_state_area),
            `Total Employment in USA` = sum(total_employment_in_state_area)
            )

data_nat_year <- data_nat_year %>% 
  mutate(`Percentage of Civilian Labor Force (USA)` =  (`Total Civilian Labor Force (USA)`/`Total non-institutional Population (USA)`)*100,
         `Percentage of Employment in USA` = (`Total Employment in USA`/`Total non-institutional Population (USA)`)*100,
         `Percentage of Unemployment in USA` = ((`Total Civilian Labor Force (USA)`-`Total Employment in USA`)/`Total non-institutional Population (USA)`)*100
    
  )



```



```{r fig.width=6, fig.asp=0.618}
data_nat_year %>%
  ggplot(aes(x=year, y = `Percentage of Unemployment in USA`)) +
  geom_line(color="steelblue") +
  geom_point() +
  theme_minimal() +
  labs(title = "Percentage of Unemployment (USA)", subtitle = "From 1976 - 2022", x = "Year", y = "Percentage of Unemployment")
```

```{r fig.width=6, fig.asp=0.618}
data_clean %>%
  ggplot(aes(x=month, y = percent_percent_of_labor_force_unemployed_in_state_area)) +
  #geom_violin(fill = "steelblue")+
  geom_boxplot(fill = "pink") +
  theme_minimal() +
  labs(title = "Percentage of Unemployment (USA)", subtitle = "From 1976 - 2022", x = "Month", y = "Percentage of Unemployment")
```

After the boxplot, we are going to eliminate outliners > 12
```{r}
data_clean <- data_clean %>% 
  filter(percent_percent_of_labor_force_unemployed_in_state_area <12)
```



```{r fig.width=6, fig.asp=0.618}
data_clean %>%
  ggplot(aes(x = percent_percent_of_labor_force_unemployed_in_state_area, y = month, fill = month)) +
  geom_density_ridges(alpha = 0.8) +
  theme_ridges() + 
  theme(legend.position = "none")
```

```{r fig.width=12, fig.height=20}
data_clean %>% 
  ggplot(aes(x= percent_percent_of_labor_force_unemployed_in_state_area, y = fct_reorder(state_area, percent_percent_of_labor_force_unemployed_in_state_area), fill = state_area)) +
  geom_density_ridges(alpha = 0.8) +
  theme_ridges() + 
  theme(legend.position = "none")
```


```{r fig.width=12, fig.height= 12}
data_clean %>%
  ggplot(aes(x = year, y = fct_reorder(state_area, percent_percent_of_labor_force_unemployed_in_state_area), fill = percent_percent_of_labor_force_unemployed_in_state_area)) +
  geom_tile_interactive(hover_nearest = TRUE) +
  scale_fill_viridis_c_interactive() +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 20))


```

```{r eval=FALSE, include=FALSE}
us_map <- us_map()

data_grouped_by_year_state_2022 <- data_grouped_by_year_state %>% 
  filter(year==2022)

us_map_new <- us_map %>% 
  full_join(data_grouped_by_year_state_2022, by = "fips")
```

```{r eval=FALSE, include=FALSE}
shape_f <- sf::st_read("s_22mr22.shp")

#Joining der Mietzins-Datensatzes (mietzins) mit Geo-Informationsdaten (shape)
#shape <- left_join(shape_f, data_grouped_by_year_state_2022, by="BFS")
```


```{r eval=FALSE, include=FALSE}
library(usmap)
library(ggplot2)



#data_grouped_by_year_state$fips <- as.numeric(data_grouped_by_year_state$fips) #convert to numeric
us_map_new %>% 
plot_usmap() + 
  scale_fill_continuous(label = scales::comma) + 
  theme(legend.position = "right")
```

```{r eval=FALSE, include=FALSE}
p <- us_map_new %>% 
  ggplot(aes(fill = `Percentage of Unemployment in USA`))+
  geom_sf_interactive(aes(data_id = state, 
                          tooltip = state, hover_nearest = TRUE 
                          , 
                      size = 3))+
  theme_minimal() +
  labs(title = "Mietzinsregionen im Kanton Z체rich", subtitle = "Klicken Sie auf Ihre Gemeinde, um an Ihre zust채ndige Durchf체hrungsstelle zu gelangen", caption = "KSA - Sozialverischerungen") +
  scale_fill_manual(values = c("#33a02c", "#1f78b4", "#b2df8a"))

# Konvertieren der ggplot-Karte in eine htmlwidgets-Objekt 
#css_default_hover <- girafe_css_bicolor(primary = "yellow", #secondary = "red")

set_girafe_defaults(
  opts_hover = opts_hover(css = ),
  opts_zoom = opts_zoom(min = 1, max = 4),
  opts_tooltip = opts_tooltip(css = "padding:3px;background-color:#333333;color:white;"),
  opts_sizing = opts_sizing(rescale = TRUE),
  opts_toolbar = opts_toolbar(saveaspng = TRUE, position = "bottom", delay_mouseout = 4000)
)

girafe(ggobj = p) %>%
  girafe_options(opts_hover(css = "fill:red;"))
```


```{r}
hist_func(data_clean, var = data_clean$percent_percent_of_labor_force_unemployed_in_state_area, 150, "Percentage of Unemployment in USA")
```



```{r}
one_way_avonva <- aov(`Percentage of Unemployment in USA` ~ state_area, data = data_grouped_by_year_state)

summary(one_way_avonva)



```

```{r include=FALSE}

data_input <- data_grouped_by_year_state %>% 
  filter(year == 2022)

data_prep <- data_input %>% 
  remove_rownames() %>% 
  column_to_rownames(var = "state_area") %>% 
  select(-year)


#scale data
data_scaled = scale(data_prep)

#f체r mixed Daten auf Basis der euklydischen Distanz
data_mixed <- daisy(data_scaled, metric='euclidean', stand = FALSE)
```


```{r}
cluster_object <- hclust(data_mixed, method = "ward.D")

plot(cluster_object, main="Baumdiagramm")
rect.hclust(as.hclust(cluster_object), k=4, border="red") # Clusteranzahl k anpassen
```
```{r}
multi.clust <- data.frame(k = 1:6) %>% group_by(k) %>% do(clust = kmeans(data_prep, .$k))
sumsq.clust <- multi.clust %>% group_by(k) %>% do(glance(.$clust[[1]]))

ggplot(sumsq.clust, aes(k, tot.withinss)) + geom_line() + geom_point()
```

```{r}
p.cluster <- data_prep %>% kmeans(., 4) #mit kmeans wird Anzahl Cluster definiert
p.cluster$cluster <- as.factor(p.cluster$cluster)

data_prep %>% 
  ggplot(aes(`Total non-institutional Population (USA)`,`Percentage of Unemployment in USA`, label = rownames(data_prep))) + 
  scale_fill_discrete(name = "Cluster") + 
  geom_label(aes(fill = p.cluster$cluster), colour = "white", 
  fontface = "bold", size=2)
```
```{r}
multi.clust <- data.frame(k = 1:6) %>% group_by(k) %>% do(clust = kmeans(data_prep, .$k))
multi.k <- multi.clust %>% group_by(k) %>% do(augment(.$clust[[1]], data_prep))

multi.k %>% 
  ggplot(aes(`Total non-institutional Population (USA)`, `Percentage of Unemployment in USA`)) + geom_point(aes(color = .cluster)) + 
  facet_wrap(~k)
```
```{r}
cluster_state <- cbind(data_prep, cluster = as.numeric(p.cluster$cluster))

cluster_state <-round(cluster_state %>% 
        group_by(cluster))
        
        
cluster_state %>%
  summarise(mean_pop = mean(`Total non-institutional Population (USA)`),
            mean_unemp = mean(`Percentage of Unemployment in USA`)
  )
            
```

```{r}
ttest <- t.test(cluster_state$`Percentage of Unemployment in USA` ~ cluster_state$cluster, var.equal= FALSE, alternative = "two.sided")

summary(one_way_avonva)
```

```{r}
one_way_avonva <- aov(`Percentage of Unemployment in USA` ~ cluster, data = cluster_state)

summary(one_way_avonva)
```
```{r}
data_input %>% 
  ggplot(aes(x=`Total non-institutional Population (USA)`,y=`Percentage of Unemployment in USA`))+
  geom_point()
```

```{r}
cor.test(data_input$`Total non-institutional Population (USA)`, data_input$`Percentage of Unemployment in USA`, method = "pearson")
```

